A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}

### 2015-03-11 CC/writeup.Rnw
additional sections I'd like

\par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these \hh{XXX write out all numbers of size ten and under in words. } 3 necessary graph attributes. \hh{XXX drive home the message that vertices, edges and layout are important, and repeat at this point} The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.  Our goal is to build off of the \texttt{ggnet} function and present an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.  
  \subsection{Data Structure}
  In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information. The vertex data frame should contain all the relevant vertex information. \emph{The only necessary variables are the vertex names/labels and the vertex degree (if undirected) or the vertex in degree and out degree (if directed).  The names and degrees are stored in columns and the rows are the vertices.  Other values of interest, such as grouping variables, for each vertex should be stored as columns in this dataframe, with an observation for each vertex.}  
  \par The edge data frame should contain all the relevant edge information.  \emph{The only necessary variables are the ``from vertex" and ``to vertex" for each edge in the network. The From and To vertices should be listed according to the names of the vertices in the vertex information data frame. Other variables may also be included for each edge, such as the edge weight or grouping variable.  As before, the variables of interest are columns in the data frame and the rows are each edge in the network.}\footnote{I'm not 100\% certain of all of this yet, but so far it seems to me to be the best way to think about networks in terms of data frames.}

@@

  \subsection{Layout Aesthetics}
  \emph{There are many layouts to choose from, and the default layout changes according the size of the graph.  For smaller networks, with less than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE. For larger networks with more than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE. **Now justify these two statements with research to be done.**} The \texttt{layout} aesthetic can also be changed to any of the layouts also found in the \texttt{igraph} package, such as Kamada-Kawai, circular, or Fructerman-Reingold. Finally, the background color of the graph can be changed with the \texttt{bgfill} aesthetic.  This aesthetic takes only the identity colors, not additional variables.\footnote{Sam's notes: I'm not sure if what I've laid out here is possible or even reasonable.  I have not yet found any research into what layout is best based on number of edges, number of vertices, density of the network, or anything else.}

>>>>

\par The \texttt{GGally} package by \citet{ggally} contains a very useful function written by Fran\c{c}ois Briatte and Moritz Marbach called \texttt{ggnet} that does allow for fairly straightforward customization of these \hh{XXX write out all numbers of size ten and under in words. } 3 necessary graph attributes. \hh{XXX drive home the message that vertices, edges and layout are important, and repeat at this point} The \texttt{ggnet} function, however, requires the graph input to be a \texttt{network} object according to the \texttt{network} package.  Our goal is to build off of the \texttt{ggnet} function and present an intuitive \texttt{geom} for network visualization within the \texttt{ggplot2} framework, without the need for objects other than simple data frames.  
%
  \subsection{Data Structure}
  In order for this geometry to work using data frames, there need to be two separate data frames given to the \texttt{geom\_net} function: one for the edge information and one for the vertex information. The vertex data frame should contain all the relevant vertex information. \emph{The only necessary variables are the vertex names/labels and the vertex degree (if undirected) or the vertex in degree and out degree (if directed).  The names and degrees are stored in columns and the rows are the vertices.  Other values of interest, such as grouping variables, for each vertex should be stored as columns in this dataframe, with an observation for each vertex.}  
  \par The edge data frame should contain all the relevant edge information.  \emph{The only necessary variables are the ``from vertex" and ``to vertex" for each edge in the network. The From and To vertices should be listed according to the names of the vertices in the vertex information data frame. Other variables may also be included for each edge, such as the edge weight or grouping variable.  As before, the variables of interest are columns in the data frame and the rows are each edge in the network.}\footnote{I'm not 100\% certain of all of this yet, but so far it seems to me to be the best way to think about networks in terms of data frames.}

@@

  \subsection{Layout Aesthetics}
  \emph{There are many layouts to choose from, and the default layout changes according the size of the graph.  For smaller networks, with less than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE. For larger networks with more than $x$ vertices and $y$ edges, the default layout is INSERT DEFAULT HERE. **Now justify these two statements with research to be done.** \hh{I don't think that you'd have to defend your choices of defaults. They are defaults, and can be changed. }} The \texttt{layout} aesthetic can also be changed to any of the layouts also found in the \texttt{igraph} package, such as Kamada-Kawai, circular, or Fruchterman-Reingold. Finally, the background color of the graph can be changed with the \texttt{bgfill} aesthetic.  This aesthetic takes only the identity colors, not additional variables.\footnote{Sam's notes: I'm not sure if what I've laid out here is possible or even reasonable.  I have not yet found any research into what layout is best based on number of edges, number of vertices, density of the network, or anything else. \hh{You're fine - those are things, which people are going to have strong opinions on without actual backup. You provide the environment for people to easily use network data in a familiar setting.}}

\section{Many Many Examples}
\section{Technical Details}
